Reanalysis Step 6: Sensitivity Analysis and E-values

### 2025-09-01 reanalysis-6-sensitivity.qmd
fix evalues

# Format E-value results
evalue_summary <- data.frame(
  Measure = c(
    "Point Estimate E-value",
    "CI Lower Bound E-value",
    "Observed OR",
    "CI Lower",
    "CI Upper"
  ),
  Value = c(
    round(evalue_result$E.values[1], 2),
    round(evalue_result$E.values[2], 2),
    round(main_or, 3),
    round(main_ci_lower, 3),
    round(main_ci_upper, 3)
  )
)

@@

# Interpretation
cat("\nInterpretation:\n")
cat(paste0(
  "An unmeasured confounder would need to have an association of at least ",
  round(evalue_result$E.values[1], 2),
  " with both lapsing and VI/blindness (on the OR scale) to fully explain the observed association.\n\n"
))
cat(paste0(
  "To shift the confidence interval to include the null, an unmeasured confounder would need an association of at least ",
  round(evalue_result$E.values[2], 2),
  " with both exposure and outcome.\n"
))
```

@@

# Create E-value plot
bias_plot <- bias_plot(evalue_result$E.values[1], evalue_result$E.values[2])

@@

```{r sensitivity-summary}
# Compile all sensitivity results
sensitivity_summary <- data.frame(
  Analysis = c(
    "Primary (Main Effects)",
    "E-value (point estimate)",
    "E-value (CI bound)",
    "Complete cases",
    "Weight truncation (max=5)",
    "Minimal adjustment",
    "Without influential obs"
  ),
  OR_Estimate = c(
    main_or,
    NA,
    NA,
    if (exists("comparison_df")) comparison_df$OR[2] else NA,
    if (exists("trunc_df")) trunc_df$OR[trunc_df$Truncation == "Max = 5"] else NA,
    if (exists("spec_df")) spec_df$OR[spec_df$Specification == "Minimal adjustment"] else NA,
    if (exists("influence_comparison")) influence_comparison$OR[2] else NA
  ),
  Interpretation = c(
    "Primary analysis result",
    paste0("Unmeasured confounding needed: OR = ", round(evalue_result$E.values[1], 2)),
    paste0("To include null: OR = ", round(evalue_result$E.values[2], 2)),
    "Missing data sensitivity",
    "Extreme weight sensitivity",
    "Model specification sensitivity",
    "Influence diagnostic"
  )
)

@@

1. **E-value**: The observed association is robust to unmeasured confounding. An unmeasured confounder would need to be strongly associated (OR ≥ `r round(evalue_result$E.values[1], 2)`) with both lapsing and VI/blindness to explain away the effect.

>>>>

# Format E-value results
# evalues.OR returns a matrix with rows "RR" and "E-values"
evalue_point <- evalue_result["E-values", "point"]
evalue_ci <- evalue_result["E-values", "lower"]

evalue_summary <- data.frame(
  Measure = c(
    "Point Estimate E-value",
    "CI Lower Bound E-value",
    "Observed OR",
    "CI Lower",
    "CI Upper"
  ),
  Value = c(
    round(evalue_point, 2),
    round(evalue_ci, 2),
    round(main_or, 3),
    round(main_ci_lower, 3),
    round(main_ci_upper, 3)
  )
)

@@

# Interpretation
cat("\nInterpretation:\n")
cat(paste0(
  "An unmeasured confounder would need to have an association of at least ",
  round(evalue_point, 2),
  " with both lapsing and VI/blindness (on the OR scale) to fully explain the observed association.\n\n"
))
cat(paste0(
  "To shift the confidence interval to include the null, an unmeasured confounder would need an association of at least ",
  round(evalue_ci, 2),
  " with both exposure and outcome.\n"
))
```

@@

# Create E-value plot
bias_plot <- bias_plot(evalue_point, evalue_ci)

@@

```{r sensitivity-summary}
# Compile all sensitivity results
sensitivity_summary <- data.frame(
  Analysis = c(
    "Primary (Main Effects)",
    "E-value (point estimate)",
    "E-value (CI bound)",
    "Complete cases",
    "Weight truncation (max=5)",
    "Minimal adjustment",
    "Without influential obs"
  ),
  OR_Estimate = c(
    main_or,
    NA,
    NA,
    if (exists("comparison_df")) comparison_df$OR[2] else NA,
    if (exists("trunc_df")) trunc_df$OR[trunc_df$Truncation == "Max = 5"] else NA,
    if (exists("spec_df")) spec_df$OR[spec_df$Specification == "Minimal adjustment"] else NA,
    if (exists("influence_comparison")) influence_comparison$OR[2] else NA
  ),
  Interpretation = c(
    "Primary analysis result",
    paste0("Unmeasured confounding needed: OR = ", round(evalue_point, 2)),
    paste0("To include null: OR = ", round(evalue_ci, 2)),
    "Missing data sensitivity",
    "Extreme weight sensitivity",
    "Model specification sensitivity",
    "Influence diagnostic"
  )
)

@@

1. **E-value**: The observed association is robust to unmeasured confounding. An unmeasured confounder would need to be strongly associated (OR ≥ `r round(evalue_point, 2)`) with both lapsing and VI/blindness to explain away the effect.
